Flat-histogram method comparison on 2D Ising model

### 2020-05-04 ising/ising-body.tex
update thesis.bib with ising bib and fix all errors in ising paper.

\davidsays{talk about phase transition, e.g. at what temperature?}

@@

All the methods except for SAMC gave a reasonable approximation for
the heat capacity peak resulting from the phase transition within 10$^9$ moves.
Figure~\ref{fig:N32-cv} shows the specific heat capacity vs. the reciprocal temperature $\beta$ at $10^{9}$ moves. Each method (except for the SAMC methods) is shown using the same single random number seed.

@@

\davidsays{Explain the range of temperature compared, and why} Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
this system. The solid/dashed lines
represent the average of the maximum value of the error in the specific heat capacity $c_V$ averaged
over eight simulations using different random number seeds. The range of maximum
errors for each simulation is shown as a shaded region. By
the time $10^8$ moves have been made all but the WL simulation have begun to
converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^9$ moves.

@@

\subsection{128 $\times$ 128 Ising system}
Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each
of the flat-histogram methods. All of the update factors initially
start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$
throughout the simulation. After about $10^{13}$ moves,
$\gamma^{\text{SAD}}_t$ proceeds as $1/t$. \davidsays{fixme similarly}
Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown. The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$. The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$. All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.

Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a function of time for
this system. The solid/dashed lines
represent the average of the maximum value of the error in the specific heat capacity $c_V$ averaged
over eight simulations using different random number seeds. The range of maximum
errors for each simulation is shown as a shaded region. By
the time $10^{13}$ moves have been made all but the WL simulation have begun to
converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$ moves.

\section{Conclusion}
\davidsays{remove this?}
The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter $\gamma_{t}^{\min}$ makes the method somewhat more difficult to implement in practice.

We find that both SAD, 1/t-WL, and WL followed by a production run
(with an adequately small $\gamma^{\min}$) demonstrate excellent and robust convergence.
They all converge more rapidly than SAMC, and unlike pure WL do not suffer from
error saturation. We find that for larger systems SAD reduces the updated factor
more slowly (and conservatively) than 1/t-WL and WL followed by a
production run. This means that SAD will take
proportionately more moves to converge to the same value as 1/t-WL as system
size is increased. SAD requires the user to specify a temperature range of
interest rather than an energy range of interest as the WL methods do. The 2D Ising model is a system where the energy range is known \emph{a priori.}, which makes a comparison between WL methods and SAD (which requires a minimum temperature) more biased towards WL methods. For the majority of cases in which the energy range of the system is not known and where a range of desired temperatures is known, this will make the SAD method
considerably more convenient. This work also visually demonstrates
that WL followed by a production run performs extremely well and is preferable to WL for ensuring both ergodicity and detailed balance.

>>>>

@@

All the methods, except for SAMC, use the same single random number seed and
give a reasonable approximation for the heat capacity peak resulting from the
phase transition within 10$^9$ moves. Figure~\ref{fig:N32-cv} shows the
specific heat capacity vs. the reciprocal temperature $\beta$ at $10^{9}$ moves.
A temperature range of $1.5k_B$ to $5k_B$ is chosen to highlight
the phase transition at the critical temperature $T_c$.

@@

Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
this system. The solid/dashed lines
represent the average of the maximum value of the error in the specific heat capacity $c_V$ averaged
over eight simulations using different random number seeds. The range of maximum
errors for each simulation is shown as a shaded region. By
the time $10^8$ moves have been made all but the WL simulation have begun to
converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^9$ moves.

@@

\subsection{128 $\times$ 128 Ising system}
For the larger 2D Ising system, the update factors initially
start at the same $\gamma_{t=0} = 1$; however, all of the methods take longer
to proceed as $1/t$ (with SAD taking $10^{13}$ moves). Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the 
flat-histogram methods. 
We implement two WL simulations followed by a production run each beginning
after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$ respectively.
Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a 
function of time for this system. The solid/dashed lines represent the average
of the maximum value of the error in the specific heat capacity $c_V$ averaged
over eight simulations using different random number seeds. The range of maximum
errors for each simulation is shown as a shaded region. By
the time $10^{13}$ moves have been made all but the WL simulation have begun to
converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$
moves. The two WL simulations followed by a production run highlight that
choosing too large a $\gamma^{\text{WL}}_t$ will result in poor convergence
properties.

\section{Conclusion}
We find that both SAD, 1/t-WL, and WL followed by a production run
(with an adequately small $\gamma^{\min}$) demonstrate excellent and robust
convergence. They all converge more rapidly than SAMC, and unlike pure WL do
not suffer from
error saturation. We find that for larger systems SAD reduces the updated factor
more slowly (and conservatively) than 1/t-WL and WL followed by a
production run. This means that SAD will take
proportionately more moves to converge to the same value as 1/t-WL as system
size is increased. SAD requires the user to specify a temperature range of
interest rather than an energy range of interest as the WL methods do. The 2D
Ising model is a system where the energy range is known \emph{a priori.}, which
makes a comparison between WL methods and SAD (which requires a minimum
temperature) more biased towards WL methods. For the majority of cases in which
the energy range of the system is not known and where a range of desired
temperatures is known, this will make the SAD method
considerably more convenient. This work also visually demonstrates
that WL followed by a production run performs extremely well and is preferable
to WL for ensuring both ergodicity and detailed balance.
